---
title: "Appendix B: Methodological Choices"
author: "Xavier Marquez"
date: today
format: gfm
execute-dir: file
execute:
  echo: false
  message: false
  warning: false
  fig-path: "figure/figs-methods-"
---

```{r setup, include=FALSE}
library(targets)
library(tidyverse)
tar_config_set(store = here::here("_targets"),
               config = "custom.yaml")

theme_set(theme_bw())

# Tuning-study tables are committed CSV snapshots (see "How this appendix is
# maintained" below), not live store objects: the store's combined tables are
# overwritten by whichever run was executed last, while these snapshots freeze
# the tuning runs' results.
tuning_perf_path <- here::here("data", "tuning_study_performance.csv")
tuning_sparsity_path <- here::here("data", "tuning_study_sparsity.csv")
has_tuning <- file.exists(tuning_perf_path) && file.exists(tuning_sparsity_path)
```

This appendix documents the main methodological choices behind the pipeline
that produces the paper's results, and reports the hyperparameter tuning study
used to select model defaults. All choices described here are recorded in the
project's configuration files (`config/corpus.yml` and `config/runs/*.yml`),
so the mapping from this text to the code is one-to-one: each subsection names
the configuration key that controls the choice it describes.

## Model hyperparameters are configuration, not code

Every predictive-model hyperparameter is set in the run configuration
(`predictive_models.engines` in `config/corpus.yml`), not in the modeling
code. An engine entry may be a bare name (which uses the documented defaults
below) or a mapping with explicit parameters; any non-default parameter set is
hashed into the affected targets' names, so changing a hyperparameter
invalidates and recomputes exactly the models it affects and nothing else.

The default hyperparameters are:

| Engine | Defaults |
|---|---|
| glmnet | `alpha = 1` (lasso), `nfolds = 10`, coefficients read at `lambda.min` |
| xgboost | `eta = 0.05`, `max_depth = 6` (set by the tuning study below; the engine-code fallback is `eta = 0.1`, `max_depth = 12`), `nrounds = 120`, `colsample_bytree = 0.5`, `subsample = 0.5`, `lambda = 0.5`, `scale_pos_weight` = class-imbalance ratio |
| LiblineaR | `type = 1` (L2-regularized L2-loss SVM, dual), `cost = 1/\bar{\lVert x \rVert}`, inverse-frequency class weights |
| naive Bayes | multinomial, `laplace = 1` |

One disclosure is in order. During this project we discovered that version 3.x
of the {xgboost} R package silently ignores the parameter list passed through
its legacy high-level interface. Any xgboost results computed with that
interface therefore reflect the library's built-in defaults
(`max_depth = 6`, `eta = 0.3`, no subsampling, no class-imbalance correction)
rather than the settings documented at the time. The pipeline now uses the
low-level `xgb.train()` interface, which honors the parameter list; all
xgboost results reported in the paper were computed after this correction.

## Hyperparameter tuning study

Defaults were checked against a grid-search tuning study run on the same
pipeline as the main results (so tuning and final models share sampling, DFM
construction, and evaluation code exactly). The study uses three widely
separated eras — volumes from 1790–1809, 1860–1879, and 1940–1959, two decade
slices each — chosen to span the large changes in corpus size, composition,
and OCR quality over the study period. In each decade slice, models were fit
on the paper's model-bearing DFM configuration (up to 500 volumes per decade,
pages restricted to those containing the target term, PPMI weighting) for
every combination in the grid:

- **glmnet:** `alpha` ∈ {0 (ridge), 0.5 (elastic net), 1 (lasso)}, plus
  `lambda.1se` as an alternative to `lambda.min` at `alpha = 1`;
- **xgboost:** `max_depth` ∈ {4, 6, 8} × `eta` ∈ {0.05, 0.1, 0.3} at
  `nrounds = 120`, plus the documented default (`max_depth = 12`,
  `eta = 0.1`) as a baseline.

Variants are compared on held-out test AUC and Cohen's kappa, and on weight
sparsity (the share of exactly-zero coefficients for glmnet; the share of
features with zero gain for xgboost), since weight sparsity propagates into
all downstream entropy and divergence measures.

```{r tuning-figures, results="asis"}
# Self-contained conditional (no cross-chunk `eval=` option, which the
# tar_quarto dependency scan cannot resolve): draw the figures if the committed
# snapshots exist, otherwise print a pending note.
if (has_tuning) {
  tuning_perf <- read_csv(tuning_perf_path, show_col_types = FALSE)
  print(
    tuning_perf %>%
      filter(.metric %in% c("roc_auc", "kap")) %>%
      ggplot(aes(x = variant, y = .estimate, color = era)) +
      geom_point(position = position_dodge(width = 0.4)) +
      facet_wrap(~.metric, scales = "free_y") +
      coord_flip() +
      labs(x = NULL, y = "Held-out estimate", color = "Era",
           title = "Tuning study: held-out performance by hyperparameter variant")
  )

  tuning_sparsity <- read_csv(tuning_sparsity_path, show_col_types = FALSE)
  print(
    tuning_sparsity %>%
      ggplot(aes(x = variant, y = zero_share, color = era)) +
      geom_point(position = position_dodge(width = 0.4)) +
      coord_flip() +
      labs(x = NULL, y = "Share of zero-weight features", color = "Era",
           title = "Tuning study: weight sparsity by hyperparameter variant")
  )
} else {
  cat("*(Tuning-study snapshot tables not yet present in `data/`; the figures",
      "will render once the tuning runs' results are snapshotted. See \"How",
      "this appendix is maintained\".)*")
}
```

**Adopted defaults.** For glmnet, the study left the defaults unchanged:
`alpha = 0.5` and `alpha = 1` are statistically indistinguishable in held-out
AUC (0.816 vs. 0.814 averaged across eras), `lambda.1se` sacrifices a little
kappa for somewhat sparser weights, and ridge (`alpha = 0`) is clearly worse
at discrimination (AUC 0.748) — so the paper keeps lasso at `lambda.min`. For
xgboost, `max_depth = 6, eta = 0.05` was adopted: it matches the best
grid cell on AUC (0.819 vs. 0.820) and kappa (0.490 vs. 0.495) within noise
while being markedly better calibrated (mean log loss 1.16 against 1.53 for
the former `max_depth = 12, eta = 0.1` setting, which ties the best AUC but
is the most overconfident of the top group). The `eta = 0.3` cells are
uniformly worst. No adopted setting sits on a worrying grid edge: performance
flattens between depth 8 and 12, and the feared undertraining of `eta = 0.05`
at 120 rounds did not materialize, so no `xgb.cv` follow-up was needed. These
defaults are recorded as engine parameters in `config/corpus.yml`, in the
same commit as this paragraph.

## From model weights to distributions

Signed model weights (regression coefficients, SVM weights, log-odds ratios,
gain shares) are converted to probability-like distributions over the
vocabulary before any entropy or divergence is computed. The conversion is
configurable (`weights.normalization`); the paper uses `positive`, which keeps
only positive weights and renormalizes. Two alternatives are implemented for
robustness checks: `shift` (shift by the minimum, then renormalize, retaining
negative-weight features at low mass) and `softmax` (exponential weighting
with a temperature). The choice matters mainly for how features that models
treat as *negative* evidence enter the entropy measures; `positive` reads the
distributions as "what the model considers evidence *for* the concept."

## Duplicate volumes are kept by default

HathiTrust contains many duplicate copies and editions of the same work. The
pipeline keeps them (`samples.dedupe: none`): a work reprinted in many copies
was more widely available, and multiplicity is precisely the popularity signal
a "corpus as readership" design should preserve. Two deduplication modes are
available as robustness checks (`bib_key`: one volume per HathiTrust
bibliographic key; `author_title`: one per author-title pair), applied within
each period slice so a reprint in a *later* decade still counts toward that
decade.

## Vocabulary handling in divergence measures

Kullback-Leibler and Jensen-Shannon divergences between period-specific weight
distributions are computed on the *union* of the two periods' vocabularies,
with additive smoothing for features absent from one side. This is the
headline measure: restricting to the intersection would silently discard
exactly the vocabulary innovation the measures are meant to detect. An
intersection-vocabulary variant can be enabled for robustness
(`kl.common_vocab_robustness: true`), which adds parallel `*_commonvocab`
targets without disturbing the headline ones.

## Evaluation uses training-set statistics

Feature weighting at evaluation time (TF-IDF document frequencies, PPMI
marginals) is computed from the *training* DFM and applied to the evaluation
DFM, never recomputed on test data — recomputing on the test set would leak
test-set statistics into the features. Out-of-distribution evaluation
("testing - OOD") applies the same rule: models trained on page-balanced
(restricted) samples are evaluated on unrestricted samples with weights
derived from their own training statistics.

## Sampling uncertainty

The headline results replicate the entire sample → DFM → train/test split →
model → weights → entropy/divergence chain **five times per decade** with
independent volume samples (`samples.n_repeats: 5`; replicate seeds are
derived deterministically from target names by the {targets} framework).
Replication is applied to the model-bearing chain (restricted samples at the
maximum volume cap); the corpus-descriptive chain is computed once. Time-series
figures report the across-replicate mean with replicate-range bands.

These are *replicate ranges*, not confidence intervals: five replicates is
far too few for tail quantiles, and no interval-coverage claim is made. The
bands answer the specific question they are designed for — whether
decade-to-decade movements in entropy and divergence exceed volume-sampling
noise — and for that purpose the effective information is larger than "n = 5"
suggests: the noise level can be pooled across the ~31 decades (over 120
degrees of freedom for the replicate variance). The replicates capture
volume-sampling, page-sampling, split, and model-fitting stochasticity; they
do not capture workset construction, OCR, or catalog uncertainty, which are
shared across replicates. Out-of-distribution evaluations pair every
replicate's models with the first replicate's unrestricted evaluation DFM.

## Reproducibility

- **Environment:** all computation runs in a versioned container
  (`ghcr.io/xmarquez/demagogues`, built from `docker/Dockerfile`) with the R
  package library pinned by `renv.lock` (R 4.5.1).
- **Configuration:** every run is fully described by a YAML file under
  `config/runs/`; the parameter grids that generate the pipeline's targets are
  derived from it deterministically.
- **Sampling:** volume sampling uses {targets}' deterministic per-target
  seeding, and the exact volumes drawn are exported: the `sampled_htids`
  target records every sampled htid × period × replicate × sample
  configuration, and is published as supplementary material so the corpus
  sample can be reconstructed exactly.

## How this appendix is maintained

This document is a `tar_quarto` target: it renders as part of the pipeline
(the `document` output group) and reads only (a) objects from the targets
store and (b) committed snapshot files under `data/`. To update it:

1. **Prose about a methodological choice** — edit this file directly; the
   relevant configuration keys are named in each section, so keep text and
   `config/corpus.yml` in sync in the same commit.
2. **Tuning-study figures** — after fetching the tuning runs' result bundles,
   snapshot their `combined_performance` and weight-sparsity summaries to
   `data/tuning_study_performance.csv` and `data/tuning_study_sparsity.csv`
   (columns used here: `era`, `variant`, `.metric`, `.estimate` /
   `zero_share`). The figures activate automatically once those files exist.
3. **Rendering** — `targets::tar_make(Appendix_Methods)` locally, or it
   renders on the cluster in any run whose `outputs` include `document`; the
   rendered markdown ships home in the results bundle.
