---
title: "Appendix"
author: "Xavier Marquez"
date: "`r date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = here::here("Paper/figure/figs-"))

library(targets)
library(tidyverse)
tar_config_set(store = here::here("_targets"),
               config = "custom.yaml")

theme_set(theme_bw())
```


## Sample Characteristics

```{r}
tar_load(usable_htids)

usable_htids %>%
  group_by(decade) %>%
  summarise(n = n_distinct(htid)) %>%
  ggplot(aes(x = decade, y = n)) +
  geom_col()
```

```{r}
tar_load(samples)
tar_load(samples_restricted)

max_vols <- samples %>%
  count(decade) %>%
  pull(n) %>%
  max()
```


Our sample includes up to `r max_vols` volumes per decade.

```{r}
samples %>%
  group_by(decade) %>%
  summarise(n = n_distinct(htid)) %>%
  ggplot(aes(x = decade, y = n)) +
  geom_col()


```
```{r}
max_vols_restricted <- samples_restricted %>%
  count(decade) %>%
  pull(n) %>%
  max()

samples_restricted %>%
  group_by(decade) %>%
  summarise(n = n_distinct(htid)) %>%
  ggplot(aes(x = decade, y = n)) +
  geom_col()
```

These represent varying proportions of the total sample, though volumes that mention democracy more often are more likely to be in the sample. (The probability of selection into the sample is weighted by the number of mentions of 'democracy' in the volume).

```{r}


per_decade <- usable_htids %>%
  group_by(decade) %>%
  summarise(total_htids = n_distinct(htid),
            total_mentions = sum(n))

samples %>%
  left_join(per_decade) %>%
  group_by(decade, total_htids) %>%
  summarise(`Proportion of volumes in decade` = n_distinct(htid)/unique(total_htids),
            `Proportion of mentions in decade` = sum(n)/unique(total_mentions)) %>%
  pivot_longer(starts_with("Proportion")) %>%
  ggplot(aes(x = decade, y = value, fill = name)) +
  geom_col(position = "dodge") +
  labs(y = "Proportion of HTIDs mentioning 'democracy'",
       fill = "") +
  scale_fill_viridis_d() +
  scale_y_continuous(labels = scales::percent)

```

```{r}
samples_restricted %>%
  left_join(per_decade) %>%
  group_by(decade, total_htids) %>%
  summarise(`Proportion of volumes in decade` = n_distinct(htid)/unique(total_htids),
            `Proportion of mentions in decade` = sum(n)/unique(total_mentions)) %>%
  pivot_longer(starts_with("Proportion")) %>%
  ggplot(aes(x = decade, y = value, fill = name)) +
  geom_col(position = "dodge") +
  labs(y = "Proportion of HTIDs mentioning 'democracy'",
       fill = "") +
  scale_fill_viridis_d() +
  scale_y_continuous(labels = scales::percent)

```


## Graphs with all weights

```{r}
tar_load(all_model_weights)
archy_acy <- all_model_weights %>% 
  filter(str_detect(word, "archy_|cracy_"), sampling_strategy == "no downsampling") %>% 
  mutate(word = str_to_lower(word))

archy_acy  %>% 
  filter(pnormed_value > 0.5) %>% 
  ggplot(aes(x = decade, y = pnormed_value)) + 
  stat_summary() + 
  geom_point(aes(color = description), alpha = 0.2) + 
  facet_wrap(~fct_reorder(word, pnormed_value)) + 
  scale_color_viridis_d() +
  theme(legend.position = "bottom")

```
```{r}
earliest_latest <- all_model_weights %>% 
  filter(sampling_strategy == "no downsampling", scaled_value > 5) %>% 
  mutate(word = str_to_lower(word)) %>%
  group_by(word) %>%
  summarise(earliest_appearance = min(decade),
            latest_appearance = max(decade),
            total_value = sum(scaled_value),
            min_value = min(scaled_value),
            max_value = max(scaled_value),
            min_pnormed = min(pnormed_value),
            max_pnormed = max(pnormed_value),
            total_appearances = n_distinct(decade),
            period = latest_appearance - earliest_appearance) %>%
  arrange(-total_value, -period, -total_appearances) %>%
  mutate(pos = str_extract(word, "_[jnvb]{2}$") %>%
           str_remove("_"),
         total_app_quantile = cut(total_appearances, breaks = 4))

terms <- earliest_latest %>%
  group_by(total_app_quantile, pos) %>%
  slice_head(n = 25) 

terms <- terms %>%
  mutate(group_id = paste(total_app_quantile, pos))

term_lists <- terms %>%
  split(terms$group_id) %>%
  map(~pull(., word)) 
```

```{r}

model_graph <- function(term_list) {
  all_model_weights %>%
    mutate(word = str_to_lower(word)) %>%
    filter(word %in% term_list, sampling_strategy == "no downsampling") %>%
    ggplot(aes(x = decade, y = pnormed_value)) +
      stat_summary(alpha =  0.4) +
      geom_point(aes(color = description), alpha = 0.2) +
      geom_smooth() +
      facet_wrap(~fct_reorder(word, pnormed_value)) +
      scale_color_viridis_d() +
      theme(legend.position = "bottom") +
      geom_hline(yintercept = 0.5, color = "red") 
}

term_lists %>%
  walk(~print(model_graph(.)))
```


## Graphs for Individual Models

```{r, child="Paper/graph_appendix.rmd"}
```

## Model performance graphs

```{r fig-combined_performance_graph_classification_auc}
#| fig-cap: "Predictive performance of classification models, AUC measure."
#| fig-width: 9
#| fig-height: 9

rm(all_model_weights)
invisible(gc())
tar_load(combined_performance)

combined_performance %>% 
  dplyr::filter(stringr::str_detect(.metric, "auc")) %>%
  dplyr::mutate(sampling_strategy = case_when(downsample ~ "random sample of majority category",
                                              TRUE ~ "no downsampling")) %>%
  ggplot2::ggplot(ggplot2::aes( x= decade, y = .estimate, color = paste(model_type), 
                                alpha = forcats::fct_rev(use),
                                linetype = sampling_strategy)) +   ggplot2::geom_point() + 
  ggplot2::geom_line() + 
  ggplot2::theme_bw() +
  ggplot2::labs(alpha = "Sample", color = "Model", shape = "Model data", linetype = "Model data",
                title = "Predictive performance of classification models, AUC measure.") +
  theme(legend.position = "bottom")+
  guides(colour = guide_legend(ncol = 1, title.position = "top"),
         alpha = guide_legend(ncol = 1, title.position = "top"),
         linetype = guide_legend(ncol = 1, title.position = "top"),
         shape = guide_legend(ncol = 1, title.position = "top"))

```

```{r fig-combined_performance_graph_classification_acc}
#| fig-cap: "Predictive performance of classification models, accuracy measure."
#| fig-width: 9
#| fig-height: 9

combined_performance %>% 
  dplyr::filter(stringr::str_detect(.metric, "accuracy")) %>%
  dplyr::mutate(sampling_strategy = case_when(downsample ~ "random sample of majority category",
                                              TRUE ~ "no downsampling")) %>%
  ggplot2::ggplot(ggplot2::aes( x= decade, y = .estimate, color = paste(model_type), 
                                alpha = forcats::fct_rev(use),
                                linetype = sampling_strategy)) + 
  ggplot2::geom_point() + 
  ggplot2::geom_line() + 
  ggplot2::theme_bw() +
  ggplot2::labs(alpha = "Sample", color = "Model", shape = "Model data", linetype = "Model data",
                title = "Predictive performance of classification models, accuracy measure.") +
  theme(legend.position = "bottom")+
  guides(colour = guide_legend(ncol = 1, title.position = "top"),
         alpha = guide_legend(ncol = 1, title.position = "top"),
         linetype = guide_legend(ncol = 1, title.position = "top"),
         shape = guide_legend(ncol = 1, title.position = "top"))
```

```{r fig-combined_performance_graph_classification_kap}
#| fig-cap: "Predictive performance of classification models, Cohen's Kappa measure."
#| fig-width: 9
#| fig-height: 9

combined_performance %>% 
  dplyr::filter(stringr::str_detect(.metric, "kap")) %>%
  dplyr::mutate(sampling_strategy = case_when(downsample ~ "random sample of majority category",
                                              TRUE ~ "no downsampling")) %>%
  ggplot2::ggplot(ggplot2::aes( x= decade, y = .estimate, color = paste(model_type), 
                                alpha = forcats::fct_rev(use),
                                linetype = sampling_strategy)) + 
  ggplot2::geom_point() + 
  ggplot2::geom_line() + 
  ggplot2::theme_bw() +
  ggplot2::labs(alpha = "Sample", color = "Model", shape = "Model data", linetype = "Model data",
                title = "Predictive performance of classification models, kap measure.") +
  theme(legend.position = "bottom")+
  guides(colour = guide_legend(ncol = 1, title.position = "top"),
         alpha = guide_legend(ncol = 1, title.position = "top"),
         linetype = guide_legend(ncol = 1, title.position = "top"),
         shape = guide_legend(ncol = 1, title.position = "top"))
```


```{r fig-combined_performance_graph_classification_log_loss}
#| fig-cap: "Predictive performance of classification models, log loss measure."
#| fig-width: 9
#| fig-height: 9

combined_performance %>% 
  dplyr::filter(stringr::str_detect(.metric, "mn_log_loss")) %>%
  dplyr::mutate(sampling_strategy = case_when(downsample ~ "random sample of majority category",
                                              TRUE ~ "no downsampling")) %>%
  ggplot2::ggplot(ggplot2::aes( x= decade, y = .estimate, color = paste(model_type), 
                                alpha = forcats::fct_rev(use),
                                linetype = sampling_strategy)) + 
  ggplot2::geom_point() + 
  ggplot2::geom_line() + 
  ggplot2::theme_bw() +
  ggplot2::labs(alpha = "Sample", color = "Model", shape = "Model data", linetype = "Model data",
                title = "Predictive performance of classification models, log loss measure.") +
  theme(legend.position = "bottom")+
  guides(colour = guide_legend(ncol = 1, title.position = "top"),
         alpha = guide_legend(ncol = 1, title.position = "top"),
         linetype = guide_legend(ncol = 1, title.position = "top"),
         shape = guide_legend(ncol = 1, title.position = "top"))
```

