---
title: 'Graphs for auth_glmnet_40'
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: right
execute-dir: project
execute:
  echo: false
  message: false
  warning: false
---

```{r setup, include=FALSE}
library(targets)
library(dplyr)
library(tidyr)
library(ggplot2)
library(here)
library(knitr)
targets::tar_config_set(store = here::here("_targets"), config = "custom.yaml")
ggplot2::theme_set(ggplot2::theme_bw())
```

Run: `auth_glmnet_40`.
Feature(s): authoritarianism.
Authoritarianism/authoritarian combo pilot: 40 restricted volumes per period, glmnet classification.

This document reads precomputed graph objects from the local targets store.

## Predictive Performance

F1 is the harmonic mean of precision and recall for pages containing the target feature. These diagnostics are internal to the configured sample and train/test split; they do not estimate prevalence in the full HathiTrust corpus.

```{r}
#| label: fig-performance-f1
#| fig-cap: "Testing F1 by model and period."
#| fig-width: 8
#| fig-height: 4
performance <- targets::tar_read_raw("combined_performance") |>
  dplyr::left_join(
    targets::tar_read_raw("info_performance") |>
      dplyr::distinct(
        performance_id,
        performance_split,
        predictive_model_engine,
        predictive_model_dfm_weight,
        sample_type,
        sample_max_vols
      ),
    by = "performance_id"
  ) |>
  dplyr::mutate(
    model = paste(
      predictive_model_engine,
      predictive_model_dfm_weight,
      "weights"
    )
  )

sample_sizes <- targets::tar_read_raw("sample_sizes") |>
  dplyr::select(
    performance_id,
    period,
    performance_split,
    sample_size,
    negative,
    positive
  )

performance_wide <- performance |>
  dplyr::filter(
    performance_split == "testing",
    .metric %in% c("f_meas", "roc_auc", "accuracy", "kap")
  ) |>
  dplyr::select(period, performance_id, model, .metric, .estimate) |>
  tidyr::pivot_wider(names_from = .metric, values_from = .estimate) |>
  dplyr::left_join(
    sample_sizes |>
      dplyr::filter(performance_split == "testing"),
    by = c("period", "performance_id")
  ) |>
  dplyr::arrange(model, period)

ggplot2::ggplot(
  performance_wide,
  ggplot2::aes(x = period, y = f_meas, color = model, group = model)
) +
  ggplot2::geom_line(linewidth = 0.6) +
  ggplot2::geom_point(size = 2) +
  ggplot2::scale_y_continuous(limits = c(0, 1)) +
  ggplot2::labs(x = "Period", y = "Testing F1", color = "Model")
```

```{r}
#| label: tbl-performance-testing
performance_wide |>
  dplyr::transmute(
    model,
    period,
    f1 = round(f_meas, 3),
    roc_auc = round(roc_auc, 3),
    accuracy = round(accuracy, 3),
    kappa = round(kap, 3),
    sample_size,
    positive,
    negative
  ) |>
  knitr::kable(
    caption = "Testing split predictive performance by model and period."
  )
```

## Term Weight Trajectories

These plots use raw predictive-model weights recovered from the targets store. Xgboost values are feature-importance scores, not signed coefficients. Facets use free y scales because model outputs are not directly comparable in magnitude. The automatic top-term plots select adjectives, verbs, and nouns separately within each model using a cumulative log-scaled positive-association score: `sum(log1p(pmax(weight, 0)))` across periods. This favors terms with persistent positive association while dampening one-period spikes.

```{r}
#| label: top-term-data
#| include: false
tracked_top_n <- 10
tracked_pos_levels <- c("Adjectives", "Verbs", "Nouns")

model_weights <- targets::tar_read_raw("predictive_model_weights") |> 
  dplyr::left_join(
    targets::tar_read_raw("info_predictive_model_weights") |> 
      dplyr::distinct(
        weight_id = weight_pred_id,
        predictive_model_engine,
        predictive_model_dfm_weight
      ),
    by = "weight_id"
  ) |> 
  dplyr::mutate(
    model = paste(
      predictive_model_engine,
      predictive_model_dfm_weight,
      "weights"
    )
  )

top_terms <- model_weights |> 
  dplyr::mutate(
    pos_group = dplyr::case_when(
      stringr::str_detect(
        word,
        stringr::regex("_jj[a-z]*$", ignore_case = TRUE)
      ) ~ "Adjectives",
      stringr::str_detect(
        word,
        stringr::regex("_vb[a-z]*$", ignore_case = TRUE)
      ) ~ "Verbs",
      stringr::str_detect(
        word,
        stringr::regex("_nn[a-z]*$", ignore_case = TRUE)
      ) ~ "Nouns",
      TRUE ~ NA_character_
    ),
    pos_group = factor(pos_group, levels = tracked_pos_levels),
    positive_value = pmax(value, 0),
    selection_score = log1p(positive_value)
  ) |> 
  dplyr::filter(!is.na(pos_group)) |> 
  dplyr::group_by(model, pos_group, word) |> 
  dplyr::summarise(
    association_score = sum(selection_score, na.rm = TRUE),
    periods_positive = sum(positive_value > 0, na.rm = TRUE),
    max_positive_weight = max(positive_value, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  dplyr::filter(association_score > 0) |> 
  dplyr::group_by(model, pos_group) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word,
    .by_group = TRUE
  ) |> 
  dplyr::slice_head(n = tracked_top_n) |> 
  dplyr::ungroup()
```

### glmnet ppmi weights - Adjectives

```{r}
#| label: fig-top-term-weights-glmnet-ppmi-weights-adjectives
#| fig-cap: "Top persistent positive-association adjective terms for glmnet ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "glmnet ppmi weights"
pos_name <- "Adjectives"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive adjective terms available for this model.")
}
```

### glmnet ppmi weights - Verbs

```{r}
#| label: fig-top-term-weights-glmnet-ppmi-weights-verbs
#| fig-cap: "Top persistent positive-association verb terms for glmnet ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "glmnet ppmi weights"
pos_name <- "Verbs"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive verb terms available for this model.")
}
```

### glmnet ppmi weights - Nouns

```{r}
#| label: fig-top-term-weights-glmnet-ppmi-weights-nouns
#| fig-cap: "Top persistent positive-association noun terms for glmnet ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "glmnet ppmi weights"
pos_name <- "Nouns"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive noun terms available for this model.")
}
```


### Selected Configured Terms

```{r}
#| label: fig-selected-term-weights
#| fig-cap: "Selected term weights or importance by model over time."
#| fig-width: 9
#| fig-height: 5
tracked_terms <- c("legitimate_jj", "rights_nn", "legitimacy_nn")

if (length(tracked_terms)) {
  selected_weights <- model_weights |> 
    dplyr::distinct(model, period) |> 
    tidyr::expand_grid(word = tracked_terms) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(word %in% tracked_terms) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = tracked_terms)
    )

  ggplot2::ggplot(
    selected_weights,
    ggplot2::aes(x = period, y = value, color = word, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
    ggplot2::geom_line(linewidth = 0.6) +
    ggplot2::geom_point(size = 1.5) +
    ggplot2::facet_wrap(ggplot2::vars(model), scales = "free_y") +
    ggplot2::labs(x = "Period", y = "Raw weight / importance", color = "Term")
} else {
  cat("No tracked terms configured.")
}
```

## Graphs


### SVD word vectors ppmi 50 dimensions - Adjectives

```{r}
#| label: fig-graph-weight-a65c1ec2
#| fig-cap: "SVD word vectors ppmi 50 dimensions; Adjectives; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_a65c1ec2")
graph
```


### SVD word vectors ppmi 50 dimensions - Nouns

```{r}
#| label: fig-graph-weight-fcb7f2e7
#| fig-cap: "SVD word vectors ppmi 50 dimensions; Nouns; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_fcb7f2e7")
graph
```


### SVD word vectors ppmi 50 dimensions - Verbs

```{r}
#| label: fig-graph-weight-c9db18d9
#| fig-cap: "SVD word vectors ppmi 50 dimensions; Verbs; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_c9db18d9")
graph
```


### SVD word vectors ppmi 50 dimensions - All terms

```{r}
#| label: fig-graph-weight-de31048b
#| fig-cap: "SVD word vectors ppmi 50 dimensions; All terms; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_de31048b")
graph
```


### Predictive model weights glmnet - Adjectives

```{r}
#| label: fig-graph-weight-f1754419
#| fig-cap: "Predictive model weights glmnet; Adjectives; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_f1754419")
graph
```


### Predictive model weights glmnet - Nouns

```{r}
#| label: fig-graph-weight-c85e14b9
#| fig-cap: "Predictive model weights glmnet; Nouns; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_c85e14b9")
graph
```


### Predictive model weights glmnet - Verbs

```{r}
#| label: fig-graph-weight-39d1a434
#| fig-cap: "Predictive model weights glmnet; Verbs; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_39d1a434")
graph
```


### Predictive model weights glmnet - All terms

```{r}
#| label: fig-graph-weight-331684c5
#| fig-cap: "Predictive model weights glmnet; All terms; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_331684c5")
graph
```


### PPMI feature_ppmi - Adjectives

```{r}
#| label: fig-graph-weight-965cb734
#| fig-cap: "PPMI feature_ppmi; Adjectives; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_965cb734")
graph
```


### PPMI feature_ppmi - Nouns

```{r}
#| label: fig-graph-weight-76747374
#| fig-cap: "PPMI feature_ppmi; Nouns; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_76747374")
graph
```


### PPMI feature_ppmi - Verbs

```{r}
#| label: fig-graph-weight-97b5b94f
#| fig-cap: "PPMI feature_ppmi; Verbs; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_97b5b94f")
graph
```


### PPMI feature_ppmi - All terms

```{r}
#| label: fig-graph-weight-08e99f58
#| fig-cap: "PPMI feature_ppmi; All terms; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_08e99f58")
graph
```


### PPMI ppmi_similarities - Adjectives

```{r}
#| label: fig-graph-weight-288b47e6
#| fig-cap: "PPMI ppmi_similarities; Adjectives; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_288b47e6")
graph
```


### PPMI ppmi_similarities - Nouns

```{r}
#| label: fig-graph-weight-2090d0a0
#| fig-cap: "PPMI ppmi_similarities; Nouns; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_2090d0a0")
graph
```


### PPMI ppmi_similarities - Verbs

```{r}
#| label: fig-graph-weight-e8ff6287
#| fig-cap: "PPMI ppmi_similarities; Verbs; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_e8ff6287")
graph
```


### PPMI ppmi_similarities - All terms

```{r}
#| label: fig-graph-weight-5ba80a99
#| fig-cap: "PPMI ppmi_similarities; All terms; 40 restricted volumes per period; lowercased. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graph <- targets::tar_read_raw("graph_weight_5ba80a99")
graph
```
