---
title: 'JSTOR graphs: authoritarianism_glmnet_100'
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: right
    include-in-header:
      text: |
        <style>
        #TOC ul.collapse {
          display: block !important;
        }
        #TOC ul.collapse li {
          margin-left: 0.75rem;
        }
        </style>
execute-dir: project
execute:
  echo: false
  message: false
  warning: false
---

```{r setup, include=FALSE}
library(targets)
library(dplyr)
library(tidyr)
library(ggplot2)
library(here)
library(knitr)
library(stringr)
source(here::here("R", "jstor_corpus_functions.R"))
targets::tar_config_set(store = here::here("_targets"), config = "custom.yaml")
ggplot2::theme_set(ggplot2::theme_bw())
```

Run: `authoritarianism_jstor_glmnet_100`.
Feature(s): authoritarianism.
Authoritarianism run over JSTOR DfR NGRAMS1: top 100 authoritarianism-positive articles per decade with matched negative articles; fits glmnet, LiblineaR, naive Bayes, and xgboost classifiers.

This document reads precomputed JSTOR graph objects from the central `_targets` store.
JSTOR DfR NGRAMS1 files are untagged unigram counts, so POS-specific Hathi panels are not generated here.

## Feature Support Audit

```{r}
#| label: tbl-jstor-feature-support
audit <- targets::tar_read(jstor_sampling_audit_authoritarianism_jstor_glmnet_100, store = here::here('_targets'))
support <- jstor_model_period_support(audit, targets::tar_read(
  jstor_config_authoritarianism_jstor_glmnet_100, store = here::here('_targets'))
)
support |>
  dplyr::mutate(
    positive_share = round(positive_texts / n_texts, 3)
  ) |>
  dplyr::select(
    period, period_label, n_texts, positive_texts, negative_texts,
    positive_share, feature_tokens, model_trainable
  ) |>
  knitr::kable(caption = "authoritarianism feature support by decade in JSTOR DfR NGRAMS1 articles.")
```

## Predictive Performance

F1 is the harmonic mean of precision and recall for JSTOR articles containing the target feature. These diagnostics are internal to the configured sample and train/test split; they do not estimate prevalence in JSTOR.

```{r}
#| label: fig-jstor-performance-f1
#| fig-cap: "Testing F1 by model and period."
#| fig-width: 8
#| fig-height: 4
performance_path <- targets::tar_read(jstor_combined_performance_authoritarianism_jstor_glmnet_100, store = here::here('_targets'))
if (!grepl("^(?:[A-Za-z]:|/|\\\\)", performance_path)) {
  performance_path <- here::here(performance_path)
}
performance <- readRDS(performance_path) |>
  dplyr::mutate(
    model = paste(
      predictive_model_engine,
      predictive_model_dfm_weight,
      "weights"
    )
  )

sample_sizes <- targets::tar_read(jstor_sample_sizes_authoritarianism_jstor_glmnet_100, store = here::here('_targets')) |>
  dplyr::select(
    performance_id,
    period,
    performance_split,
    sample_size,
    negative,
    positive
  )

performance_wide <- performance |>
  dplyr::filter(
    performance_split == "testing",
    .metric %in% c("f_meas", "roc_auc", "accuracy", "kap")
  ) |>
  dplyr::select(period, performance_id, model, .metric, .estimate) |>
  tidyr::pivot_wider(names_from = .metric, values_from = .estimate) |>
  dplyr::left_join(
    sample_sizes |>
      dplyr::filter(performance_split == "testing"),
    by = c("period", "performance_id")
  ) |>
  dplyr::arrange(model, period)

ggplot2::ggplot(
  performance_wide,
  ggplot2::aes(x = period, y = f_meas, color = model, group = model)
) +
  ggplot2::geom_line(linewidth = 0.6) +
  ggplot2::geom_point(size = 2) +
  ggplot2::scale_y_continuous(limits = c(0, 1)) +
  ggplot2::labs(x = "Period", y = "Testing F1", color = "Model")
```

```{r}
#| label: tbl-jstor-performance-testing
performance_wide |>
  dplyr::transmute(
    model,
    period,
    f1 = round(f_meas, 3),
    roc_auc = round(roc_auc, 3),
    accuracy = round(accuracy, 3),
    kappa = round(kap, 3),
    sample_size,
    positive,
    negative
  ) |>
  knitr::kable(
    caption = "Testing split predictive performance by model and period."
  )
```

## Term Weight Trajectories

These plots use raw predictive-model weights recovered from the targets store. Xgboost values are feature-importance scores, not signed coefficients. Facets use free y scales because model outputs are not directly comparable in magnitude. The automatic top-term plots select terms using a cumulative log-scaled positive-association score: `sum(log1p(pmax(weight, 0)))` across periods.

```{r}
#| label: jstor-top-term-data
#| include: false
tracked_top_n <- 10

model_weights <- targets::tar_read(jstor_predictive_model_weights_authoritarianism_jstor_glmnet_100, store = here::here('_targets')) |> 
  dplyr::mutate(
    model = paste(
      predictive_model_engine,
      predictive_model_dfm_weight,
      "weights"
    )
  )

top_terms <- model_weights |> 
  dplyr::mutate(
    positive_value = pmax(value, 0),
    selection_score = log1p(positive_value)
  ) |> 
  dplyr::group_by(model, word) |> 
  dplyr::summarise(
    association_score = sum(selection_score, na.rm = TRUE),
    periods_positive = sum(positive_value > 0, na.rm = TRUE),
    max_positive_weight = max(positive_value, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  dplyr::filter(association_score > 0) |> 
  dplyr::group_by(model) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word,
    .by_group = TRUE
  ) |> 
  dplyr::slice_head(n = tracked_top_n) |> 
  dplyr::ungroup()
```
### glmnet ppmi weights - All Terms

```{r}
#| label: fig-jstor-top-term-weights-glmnet-ppmi-weights-all-terms
#| fig-cap: "Top persistent positive-association terms for glmnet ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "glmnet ppmi weights"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = "Terms ordered by cumulative log-scaled positive association"
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive terms available for this model.")
}
```

### LiblineaR ppmi weights - All Terms

```{r}
#| label: fig-jstor-top-term-weights-liblinear-ppmi-weights-all-terms
#| fig-cap: "Top persistent positive-association terms for LiblineaR ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "LiblineaR ppmi weights"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = "Terms ordered by cumulative log-scaled positive association"
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive terms available for this model.")
}
```

### naivebayes ppmi weights - All Terms

```{r}
#| label: fig-jstor-top-term-weights-naivebayes-ppmi-weights-all-terms
#| fig-cap: "Top persistent positive-association terms for naivebayes ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "naivebayes ppmi weights"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = "Terms ordered by cumulative log-scaled positive association"
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive terms available for this model.")
}
```

### xgboost ppmi weights - All Terms

```{r}
#| label: fig-jstor-top-term-weights-xgboost-ppmi-weights-all-terms
#| fig-cap: "Top persistent positive-association terms for xgboost ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "xgboost ppmi weights"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = "Terms ordered by cumulative log-scaled positive association"
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive terms available for this model.")
}
```


### Selected Configured Terms

```{r}
#| label: fig-jstor-selected-term-weights
#| fig-cap: "Selected term weights or importance by model over time."
#| fig-width: 9
#| fig-height: 5
tracked_terms <- c("legitimate", "rights", "legitimacy")

if (length(tracked_terms)) {
  selected_weights <- model_weights |> 
    dplyr::distinct(model, period) |> 
    tidyr::expand_grid(word = tracked_terms) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(word %in% tracked_terms) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = tracked_terms)
    )

  ggplot2::ggplot(
    selected_weights,
    ggplot2::aes(x = period, y = value, color = word, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
    ggplot2::geom_line(linewidth = 0.6) +
    ggplot2::geom_point(size = 1.5) +
    ggplot2::facet_wrap(ggplot2::vars(model), scales = "free_y") +
    ggplot2::labs(x = "Period", y = "Raw weight / importance", color = "Term")
} else {
  cat("No tracked terms configured.")
}
```

## Graphs


### Predictive model weights glmnet - All terms

```{r}
#| label: fig-jstor-jstor-authoritarianism-jstor-glmnet-100-glmnet-
#| fig-cap: "Predictive model weights glmnet; All terms; top 100 feature-positive JSTOR articles per period with matched negative articles. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(jstor_graphs_authoritarianism_jstor_glmnet_100, store = here::here('_targets'))
graph_row <- graphs |> dplyr::filter(graph_id == "jstor_authoritarianism_jstor_glmnet_100_glmnet__")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights LiblineaR - All terms

```{r}
#| label: fig-jstor-jstor-authoritarianism-jstor-glmnet-100-LiblineaR-
#| fig-cap: "Predictive model weights LiblineaR; All terms; top 100 feature-positive JSTOR articles per period with matched negative articles. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(jstor_graphs_authoritarianism_jstor_glmnet_100, store = here::here('_targets'))
graph_row <- graphs |> dplyr::filter(graph_id == "jstor_authoritarianism_jstor_glmnet_100_LiblineaR__")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights naivebayes - All terms

```{r}
#| label: fig-jstor-jstor-authoritarianism-jstor-glmnet-100-naivebayes-
#| fig-cap: "Predictive model weights naivebayes; All terms; top 100 feature-positive JSTOR articles per period with matched negative articles. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(jstor_graphs_authoritarianism_jstor_glmnet_100, store = here::here('_targets'))
graph_row <- graphs |> dplyr::filter(graph_id == "jstor_authoritarianism_jstor_glmnet_100_naivebayes__")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights xgboost - All terms

```{r}
#| label: fig-jstor-jstor-authoritarianism-jstor-glmnet-100-xgboost-
#| fig-cap: "Predictive model weights xgboost; All terms; top 100 feature-positive JSTOR articles per period with matched negative articles. Terms at the top are more closely associated with 'authoritarianism' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(jstor_graphs_authoritarianism_jstor_glmnet_100, store = here::here('_targets'))
graph_row <- graphs |> dplyr::filter(graph_id == "jstor_authoritarianism_jstor_glmnet_100_xgboost__")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```
