---
title: "Latin Perseus Report: auctoritas"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: right
    df-print: paged
execute:
  warning: false
  message: false
---

```{r setup}
library(dplyr)
library(ggplot2)
library(knitr)
library(stringr)
library(targets)
library(here)
targets::tar_config_set(store = here::here("_targets"), config = "custom.yaml")
theme_set(theme_minimal(base_size = 12))
```

Initial Perseus-only Latin run for auctoritas. This uses occurrence-centered windows around configured forms and reports context-term PPMI, SVD dimensions, and broad period/author summaries.

## Occurrences

```{r occurrence-summary}
summary <- tar_read(latin_window_summary)
if (nrow(summary)) {
  summary %>%
    arrange(desc(n_occurrences)) %>%
    head(30) %>%
    knitr::kable()
} else {
  cat('No occurrences found for this run.')
}
```

## PPMI Associations

### Overall

```{r global_ppmi_terms}
#| fig-height: 7
#| fig-width: 9
global_ppmi <- tar_read(latin_global_ppmi_weights)
if (nrow(global_ppmi)) {
  plot_df <- global_ppmi %>%
    slice_max(value, n = 30, with_ties = FALSE) %>%
    mutate(word = reorder(word, value))
  ggplot(plot_df, aes(value, word)) +
    geom_col(fill = '#2f5f6a') +
    labs(x = 'PPMI x log count', y = NULL)
} else {
  cat('No global PPMI associations available.')
}
```

### By Period

```{r period_ppmi_terms}
#| fig-height: 8
#| fig-width: 10
period_ppmi <- tar_read(latin_period_ppmi_weights)
if (nrow(period_ppmi)) {
  label_col <- if ('period_label' %in% names(period_ppmi)) 'period_label' else 'period'
  plot_df <- period_ppmi %>%
    group_by(.data[[label_col]]) %>%
    slice_max(value, n = 10, with_ties = FALSE) %>%
    ungroup() %>%
    mutate(word = tidytext::reorder_within(word, value, .data[[label_col]]))
  ggplot(plot_df, aes(value, word)) +
    geom_col(fill = '#2f5f6a') +
    facet_wrap(vars(.data[[label_col]]), scales = 'free_y') +
    tidytext::scale_y_reordered() +
    labs(x = 'PPMI x log count', y = NULL)
} else {
  cat('No period PPMI associations available.')
}
```

## SVD Context Dimensions

### Term Loadings

```{r svd_term_loadings}
#| fig-height: 10
#| fig-width: 10
svd_terms <- tar_read(latin_svd_terms)
if (nrow(svd_terms)) {
  plot_df <- svd_terms %>%
    filter(dimension <= 5) %>%
    mutate(
      panel = paste('Dimension', dimension, side),
      signed_loading = if_else(side == 'negative', -abs_loading, abs_loading),
      word = tidytext::reorder_within(word, abs_loading, panel)
    )
  ggplot(plot_df, aes(abs_loading, word, fill = side)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(vars(panel), scales = 'free_y', ncol = 2) +
    tidytext::scale_y_reordered() +
    scale_fill_manual(values = c(negative = '#8a4f45', positive = '#2f5f6a')) +
    labs(x = 'Absolute loading', y = NULL)
} else {
  cat('No SVD term loadings available.')
}
```

### Window Map

```{r svd_window_map}
#| fig-height: 6
#| fig-width: 8
svd_windows <- tar_read(latin_svd_windows)
if (all(c('dim_1', 'dim_2') %in% names(svd_windows))) {
  label_col <- if ('period_label' %in% names(svd_windows)) 'period_label' else 'period'
  centroids <- svd_windows %>%
    group_by(.data[[label_col]]) %>%
    summarise(dim_1 = median(dim_1, na.rm = TRUE), dim_2 = median(dim_2, na.rm = TRUE), .groups = 'drop')
  ggplot(svd_windows, aes(dim_1, dim_2, color = .data[[label_col]])) +
    geom_point(alpha = 0.35, size = 1.2) +
    geom_text(data = centroids, aes(label = .data[[label_col]]), color = 'black', size = 3.5) +
    labs(x = 'SVD dimension 1', y = 'SVD dimension 2', color = NULL)
} else {
  cat('SVD did not produce at least two dimensions.')
}
```

## Period-Distinctive Context Terms

```{r top_period_terms}
#| fig-height: 7
#| fig-width: 10
period_terms <- tar_read(latin_period_term_weights)
if (nrow(period_terms)) {
  label_col <- if ('period_label' %in% names(period_terms)) 'period_label' else 'period'
  plot_df <- period_terms %>%
    group_by(.data[[label_col]]) %>%
    slice_max(value, n = 12, with_ties = FALSE) %>%
    ungroup() %>%
    mutate(word = tidytext::reorder_within(word, value, .data[[label_col]]))
  ggplot(plot_df, aes(value, word)) +
    geom_col(fill = '#2f5f6a') +
    facet_wrap(vars(.data[[label_col]]), scales = 'free_y') +
    tidytext::scale_y_reordered() +
    labs(x = 'Context weight', y = NULL)
} else {
  cat('No period term weights available.')
}
```

## Top Context Terms By Author

```{r top_author_terms}
#| fig-height: 10
#| fig-width: 10
author_terms <- tar_read(latin_author_term_weights)
if (nrow(author_terms)) {
  top_authors <- tar_read(latin_window_summary) %>%
    group_by(author) %>%
    summarise(n_occurrences = sum(n_occurrences), .groups = 'drop') %>%
    slice_max(n_occurrences, n = 8, with_ties = FALSE) %>%
    pull(author)
  plot_df <- author_terms %>%
    filter(author %in% top_authors) %>%
    group_by(author) %>%
    slice_max(value, n = 8, with_ties = FALSE) %>%
    ungroup() %>%
    mutate(word = tidytext::reorder_within(word, value, author))
  ggplot(plot_df, aes(value, word)) +
    geom_col(fill = '#6f5d2f') +
    facet_wrap(vars(author), scales = 'free_y') +
    tidytext::scale_y_reordered() +
    labs(x = 'Context weight', y = NULL)
} else {
  cat('No author term weights available.')
}
```

## Tracked Terms

```{r tracked_terms}
#| fig-height: 5
#| fig-width: 9
tracked <- tar_read(latin_tracked_terms_period)
if (nrow(tracked)) {
  label_col <- if ('period_label' %in% names(tracked)) 'period_label' else 'period'
  ggplot(tracked, aes(.data[[label_col]], rate_per_window, group = word)) +
    geom_line(color = '#2f5f6a') +
    geom_point(color = '#2f5f6a') +
    facet_wrap(vars(word), scales = 'free_y') +
    labs(x = NULL, y = 'Uses per occurrence window')
} else {
  cat('No configured tracked terms were found in the period weights.')
}
```

## Concordance Sample

```{r concordance-sample}
windows <- tar_read(latin_windows)
if (nrow(windows)) {
  windows %>%
    select(author, title, matched_form, concordance) %>%
    head(30) %>%
    knitr::kable()
} else {
  cat('No concordance lines available.')
}
```
