---
title: 'Graphs for authority_marxism_glmnet_100'
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: right
    include-in-header:
      text: |
        <style>
        #TOC ul.collapse {
          display: block !important;
        }
        #TOC ul.collapse li {
          margin-left: 0.75rem;
        }
        </style>
execute-dir: project
execute:
  echo: false
  message: false
  warning: false
---

```{r setup, include=FALSE}
library(targets)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(stringr)
ggplot2::theme_set(ggplot2::theme_bw())
```

Run: `authority_marxism_glmnet_100`.
Feature(s): authority.
Authority run over the annotated Marx/Engels corpus: up to 100 positive segments per decade with matched negative segments; fits glmnet, LiblineaR, naive Bayes, and xgboost classifiers.

This document reads precomputed Marx/Engels graph objects from the central `_targets` store.

## Feature Support Audit

The audit table reports segment-level support before model fitting. A segment is a fixed group of CoreNLP sentences from the annotated Marx/Engels corpus.

```{r}
#| label: tbl-marxism-feature-support
audit <- targets::tar_read(marxism_feature_audit, store = '_targets')
audit$segment_support |>
  dplyr::mutate(
    positive_share = round(positive_segments / n_segments, 3)
  ) |>
  dplyr::select(
    period, period_label, n_segments, positive_segments, negative_segments,
    positive_share, n_texts, positive_texts, trainable
  ) |>
  knitr::kable(caption = "Authority feature support by decade in Marx/Engels segments.")
```

## Predictive Performance

F1 is the harmonic mean of precision and recall for segments containing the target feature. These diagnostics are internal to the configured sample and train/test split; they do not estimate prevalence in the full Marx/Engels corpus.

```{r}
#| label: fig-marxism-performance-f1
#| fig-cap: "Testing F1 by model and period."
#| fig-width: 8
#| fig-height: 4
performance <- readRDS(targets::tar_read(marxism_combined_performance, store = '_targets')) |>
  dplyr::mutate(
    model = paste(
      predictive_model_engine,
      predictive_model_dfm_weight,
      "weights"
    )
  )

sample_sizes <- targets::tar_read(marxism_sample_sizes, store = '_targets') |>
  dplyr::select(
    performance_id,
    period,
    performance_split,
    sample_size,
    negative,
    positive
  )

performance_wide <- performance |>
  dplyr::filter(
    performance_split == "testing",
    .metric %in% c("f_meas", "roc_auc", "accuracy", "kap")
  ) |>
  dplyr::select(period, performance_id, model, .metric, .estimate) |>
  tidyr::pivot_wider(names_from = .metric, values_from = .estimate) |>
  dplyr::left_join(
    sample_sizes |>
      dplyr::filter(performance_split == "testing"),
    by = c("period", "performance_id")
  ) |>
  dplyr::arrange(model, period)

ggplot2::ggplot(
  performance_wide,
  ggplot2::aes(x = period, y = f_meas, color = model, group = model)
) +
  ggplot2::geom_line(linewidth = 0.6) +
  ggplot2::geom_point(size = 2) +
  ggplot2::scale_y_continuous(limits = c(0, 1)) +
  ggplot2::labs(x = "Period", y = "Testing F1", color = "Model")
```

```{r}
#| label: tbl-marxism-performance-testing
performance_wide |>
  dplyr::transmute(
    model,
    period,
    f1 = round(f_meas, 3),
    roc_auc = round(roc_auc, 3),
    accuracy = round(accuracy, 3),
    kappa = round(kap, 3),
    sample_size,
    positive,
    negative
  ) |>
  knitr::kable(
    caption = "Testing split predictive performance by model and period."
  )
```

## Term Weight Trajectories

These plots use raw predictive-model weights recovered from the targets store. Xgboost values are feature-importance scores, not signed coefficients. Facets use free y scales because model outputs are not directly comparable in magnitude. The automatic top-term plots select adjectives, verbs, and nouns separately within each model using a cumulative log-scaled positive-association score: `sum(log1p(pmax(weight, 0)))` across periods. This favors terms with persistent positive association while dampening one-period spikes.

```{r}
#| label: marxism-top-term-data
#| include: false
tracked_top_n <- 10
tracked_pos_levels <- c("Adjectives", "Verbs", "Nouns")

model_weights <- targets::tar_read(marxism_predictive_model_weights, store = '_targets') |> 
  dplyr::mutate(
    model = paste(
      predictive_model_engine,
      predictive_model_dfm_weight,
      "weights"
    )
  )

top_terms <- model_weights |> 
  dplyr::mutate(
    pos_group = dplyr::case_when(
      stringr::str_detect(
        word,
        stringr::regex("_jj[a-z]*$", ignore_case = TRUE)
      ) ~ "Adjectives",
      stringr::str_detect(
        word,
        stringr::regex("_vb[a-z]*$", ignore_case = TRUE)
      ) ~ "Verbs",
      stringr::str_detect(
        word,
        stringr::regex("_nn[a-z]*$", ignore_case = TRUE)
      ) ~ "Nouns",
      TRUE ~ NA_character_
    ),
    pos_group = factor(pos_group, levels = tracked_pos_levels),
    positive_value = pmax(value, 0),
    selection_score = log1p(positive_value)
  ) |> 
  dplyr::filter(!is.na(pos_group)) |> 
  dplyr::group_by(model, pos_group, word) |> 
  dplyr::summarise(
    association_score = sum(selection_score, na.rm = TRUE),
    periods_positive = sum(positive_value > 0, na.rm = TRUE),
    max_positive_weight = max(positive_value, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  dplyr::filter(association_score > 0) |> 
  dplyr::group_by(model, pos_group) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word,
    .by_group = TRUE
  ) |> 
  dplyr::slice_head(n = tracked_top_n) |> 
  dplyr::ungroup()
```
### glmnet ppmi weights - Adjectives

```{r}
#| label: fig-marxism-top-term-weights-glmnet-ppmi-weights-adjectives
#| fig-cap: "Top persistent positive-association adjective terms for glmnet ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "glmnet ppmi weights"
pos_name <- "Adjectives"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive adjective terms available for this model.")
}
```

### glmnet ppmi weights - Verbs

```{r}
#| label: fig-marxism-top-term-weights-glmnet-ppmi-weights-verbs
#| fig-cap: "Top persistent positive-association verb terms for glmnet ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "glmnet ppmi weights"
pos_name <- "Verbs"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive verb terms available for this model.")
}
```

### glmnet ppmi weights - Nouns

```{r}
#| label: fig-marxism-top-term-weights-glmnet-ppmi-weights-nouns
#| fig-cap: "Top persistent positive-association noun terms for glmnet ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "glmnet ppmi weights"
pos_name <- "Nouns"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive noun terms available for this model.")
}
```

### LiblineaR ppmi weights - Adjectives

```{r}
#| label: fig-marxism-top-term-weights-liblinear-ppmi-weights-adjectives
#| fig-cap: "Top persistent positive-association adjective terms for LiblineaR ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "LiblineaR ppmi weights"
pos_name <- "Adjectives"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive adjective terms available for this model.")
}
```

### LiblineaR ppmi weights - Verbs

```{r}
#| label: fig-marxism-top-term-weights-liblinear-ppmi-weights-verbs
#| fig-cap: "Top persistent positive-association verb terms for LiblineaR ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "LiblineaR ppmi weights"
pos_name <- "Verbs"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive verb terms available for this model.")
}
```

### LiblineaR ppmi weights - Nouns

```{r}
#| label: fig-marxism-top-term-weights-liblinear-ppmi-weights-nouns
#| fig-cap: "Top persistent positive-association noun terms for LiblineaR ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "LiblineaR ppmi weights"
pos_name <- "Nouns"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive noun terms available for this model.")
}
```

### naivebayes ppmi weights - Adjectives

```{r}
#| label: fig-marxism-top-term-weights-naivebayes-ppmi-weights-adjectives
#| fig-cap: "Top persistent positive-association adjective terms for naivebayes ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "naivebayes ppmi weights"
pos_name <- "Adjectives"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive adjective terms available for this model.")
}
```

### naivebayes ppmi weights - Verbs

```{r}
#| label: fig-marxism-top-term-weights-naivebayes-ppmi-weights-verbs
#| fig-cap: "Top persistent positive-association verb terms for naivebayes ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "naivebayes ppmi weights"
pos_name <- "Verbs"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive verb terms available for this model.")
}
```

### naivebayes ppmi weights - Nouns

```{r}
#| label: fig-marxism-top-term-weights-naivebayes-ppmi-weights-nouns
#| fig-cap: "Top persistent positive-association noun terms for naivebayes ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "naivebayes ppmi weights"
pos_name <- "Nouns"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive noun terms available for this model.")
}
```

### xgboost ppmi weights - Adjectives

```{r}
#| label: fig-marxism-top-term-weights-xgboost-ppmi-weights-adjectives
#| fig-cap: "Top persistent positive-association adjective terms for xgboost ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "xgboost ppmi weights"
pos_name <- "Adjectives"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive adjective terms available for this model.")
}
```

### xgboost ppmi weights - Verbs

```{r}
#| label: fig-marxism-top-term-weights-xgboost-ppmi-weights-verbs
#| fig-cap: "Top persistent positive-association verb terms for xgboost ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "xgboost ppmi weights"
pos_name <- "Verbs"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive verb terms available for this model.")
}
```

### xgboost ppmi weights - Nouns

```{r}
#| label: fig-marxism-top-term-weights-xgboost-ppmi-weights-nouns
#| fig-cap: "Top persistent positive-association noun terms for xgboost ppmi weights."
#| fig-width: 9
#| fig-height: 10
model_name <- "xgboost ppmi weights"
pos_name <- "Nouns"

model_top_terms <- top_terms |> 
  dplyr::filter(model == model_name, pos_group == pos_name) |> 
  dplyr::arrange(
    dplyr::desc(association_score),
    dplyr::desc(periods_positive),
    dplyr::desc(max_positive_weight),
    word
  )

if (nrow(model_top_terms)) {
  model_top_weights <- tidyr::expand_grid(
    model = model_name,
    period = sort(unique(model_weights$period[model_weights$model == model_name])),
    word = model_top_terms$word
  ) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(model == model_name) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = model_top_terms$word)
    )

  ggplot2::ggplot(
    model_top_weights,
    ggplot2::aes(x = period, y = value, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey75", linewidth = 0.3) +
    ggplot2::geom_line(color = "#2F5D62", linewidth = 0.65) +
    ggplot2::geom_point(color = "#2F5D62", size = 1.3) +
    ggplot2::facet_wrap(ggplot2::vars(word), ncol = 1, scales = "free_y") +
    ggplot2::labs(
      x = "Period",
      y = "Raw weight / importance",
      subtitle = paste(
        "Terms ordered by cumulative log-scaled positive association within",
        tolower(pos_name)
      )
    ) +
    ggplot2::theme(
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold")
    )
} else {
  cat("No positive noun terms available for this model.")
}
```


### Selected Configured Terms

```{r}
#| label: fig-marxism-selected-term-weights
#| fig-cap: "Selected term weights or importance by model over time."
#| fig-width: 9
#| fig-height: 5
tracked_terms <- c("legitimate_jj", "rights_nn", "legitimacy_nn")

if (length(tracked_terms)) {
  selected_weights <- model_weights |> 
    dplyr::distinct(model, period) |> 
    tidyr::expand_grid(word = tracked_terms) |> 
    dplyr::left_join(
      model_weights |> 
        dplyr::filter(word %in% tracked_terms) |> 
        dplyr::select(model, period, word, value),
      by = c("model", "period", "word")
    ) |> 
    dplyr::mutate(
      value = tidyr::replace_na(value, 0),
      word = factor(word, levels = tracked_terms)
    )

  ggplot2::ggplot(
    selected_weights,
    ggplot2::aes(x = period, y = value, color = word, group = word)
  ) +
    ggplot2::geom_hline(yintercept = 0, color = "grey70", linewidth = 0.3) +
    ggplot2::geom_line(linewidth = 0.6) +
    ggplot2::geom_point(size = 1.5) +
    ggplot2::facet_wrap(ggplot2::vars(model), scales = "free_y") +
    ggplot2::labs(x = "Period", y = "Raw weight / importance", color = "Term")
} else {
  cat("No tracked terms configured.")
}
```

## Graphs


### Predictive model weights glmnet - All terms

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-glmnet-
#| fig-cap: "Predictive model weights glmnet; All terms; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_glmnet__")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights glmnet - Nouns

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-glmnet-nn-NN-
#| fig-cap: "Predictive model weights glmnet; Nouns; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_glmnet__nn_NN_")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights glmnet - Verbs

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-glmnet-vb-VB-
#| fig-cap: "Predictive model weights glmnet; Verbs; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_glmnet__vb_VB_")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights glmnet - Adjectives

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-glmnet-jj-JJ-
#| fig-cap: "Predictive model weights glmnet; Adjectives; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_glmnet__jj_JJ_")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights LiblineaR - All terms

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-LiblineaR-
#| fig-cap: "Predictive model weights LiblineaR; All terms; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_LiblineaR__")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights LiblineaR - Nouns

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-LiblineaR-nn-NN-
#| fig-cap: "Predictive model weights LiblineaR; Nouns; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_LiblineaR__nn_NN_")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights LiblineaR - Verbs

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-LiblineaR-vb-VB-
#| fig-cap: "Predictive model weights LiblineaR; Verbs; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_LiblineaR__vb_VB_")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights LiblineaR - Adjectives

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-LiblineaR-jj-JJ-
#| fig-cap: "Predictive model weights LiblineaR; Adjectives; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_LiblineaR__jj_JJ_")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights naivebayes - All terms

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-naivebayes-
#| fig-cap: "Predictive model weights naivebayes; All terms; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_naivebayes__")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights naivebayes - Nouns

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-naivebayes-nn-NN-
#| fig-cap: "Predictive model weights naivebayes; Nouns; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_naivebayes__nn_NN_")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights naivebayes - Verbs

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-naivebayes-vb-VB-
#| fig-cap: "Predictive model weights naivebayes; Verbs; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_naivebayes__vb_VB_")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights naivebayes - Adjectives

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-naivebayes-jj-JJ-
#| fig-cap: "Predictive model weights naivebayes; Adjectives; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_naivebayes__jj_JJ_")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights xgboost - All terms

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-xgboost-
#| fig-cap: "Predictive model weights xgboost; All terms; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_xgboost__")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights xgboost - Nouns

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-xgboost-nn-NN-
#| fig-cap: "Predictive model weights xgboost; Nouns; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_xgboost__nn_NN_")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights xgboost - Verbs

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-xgboost-vb-VB-
#| fig-cap: "Predictive model weights xgboost; Verbs; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_xgboost__vb_VB_")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```


### Predictive model weights xgboost - Adjectives

```{r}
#| label: fig-marxism-marxism-authority-marxism-glmnet-100-xgboost-jj-JJ-
#| fig-cap: "Predictive model weights xgboost; Adjectives; up to 100 positive segments per period with matched negative segments. Terms at the top are more closely associated with 'authority' in later years. The label marks the peak density of association."
#| fig-width: 11
#| fig-height: 9
graphs <- targets::tar_read(marxism_graphs, store = '_targets')
graph_row <- graphs |> dplyr::filter(graph_id == "marxism_authority_marxism_glmnet_100_xgboost__jj_JJ_")
if (nrow(graph_row)) {
  graph_row$plot[[1]]
} else {
  cat("Graph not available.")
}
```
