---
title: "Prototype: ism-term clustering across models and repeats"
subtitle: "Democracy full run - diagnostic from saved targets"
format:
  html:
    toc: true
    toc-location: right
    fig-width: 10
    fig-height: 6
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(purrr)
library(tibble)
library(knitr)
library(here)
library(targets)

theme_set(theme_bw(base_size = 12))

store <- here::here("_targets")
targets::tar_config_set(store = store, config = "custom.yaml")
rd <- function(n) targets::tar_read_raw(n, store = store)

model_order <- c(
  "glmnet",
  "LiblineaR",
  "naivebayes",
  "xgboost",
  "PPMI feature",
  "PPMI similarities",
  "SVD PPMI 50d"
)

model_pal <- c(
  glmnet = "#1b9e77",
  LiblineaR = "#d95f02",
  naivebayes = "#7570b3",
  xgboost = "#e7298a",
  "PPMI feature" = "#66a61e",
  "PPMI similarities" = "#a6761d",
  "SVD PPMI 50d" = "#1f78b4"
)
```

## Aim

This prototype asks which `ism_` terms are stable across the five sample repeats
and across model families. It reads only saved targets from `_targets/`; it does
not rerun the pipeline.

The plots collapse part-of-speech variants, so `despotism_nn`,
`despotism_nnp`, and similar variants contribute to one `despotism` row. Within
each model/repeat source, term scores are normalized before comparing across
models because glmnet, LiblineaR, naive Bayes, xgboost, PPMI, and SVD weights
are not on a common scale.

```{r data}
info <- rd("info_all_model_weights") |>
  mutate(
    weight_id = coalesce(weight_pred_id, weight_ppmi_id, weight_svd_id),
    model_family = case_when(
      predictive_model_engine %in% c("glmnet", "LiblineaR", "naivebayes", "xgboost") ~
        predictive_model_engine,
      ppmi_fun == "feature_ppmi" ~ "PPMI feature",
      ppmi_fun == "ppmi_similarities" ~ "PPMI similarities",
      !is.na(svd_dims) ~ paste("SVD", toupper(svd_weight), paste0(svd_dims, "d")),
      TRUE ~ predictive_model_engine
    ),
    model_family = factor(model_family, levels = model_order),
    source_id = paste0(as.character(model_family), " r", sample_rep)
  ) |>
  select(weight_id, model_family, sample_rep, source_id)

weights_all <- rd("all_model_weights")
ism_weights <- weights_all[
  grepl("ism_", weights_all$word, fixed = TRUE),
  c("weight_id", "word", "period", "value")
]
rm(weights_all)
gc()

ism_weights <- ism_weights |>
  as_tibble() |>
  inner_join(info, by = "weight_id") |>
  mutate(
    term = str_remove(word, "_[a-z]+$"),
    value = as.numeric(value),
    positive_value = pmax(value, 0)
  ) |>
  group_by(weight_id) |>
  mutate(
    source_max = max(positive_value, na.rm = TRUE),
    value_norm = if_else(source_max > 0, positive_value / source_max, 0)
  ) |>
  ungroup()

diagnostics <- tibble(
  ism_rows = nrow(ism_weights),
  unique_terms = n_distinct(ism_weights$term),
  sources = n_distinct(ism_weights$weight_id),
  model_families = n_distinct(ism_weights$model_family)
)
```

```{r tbl-diagnostics}
kable(diagnostics, caption = "Data used for the ism-term prototype.")
```

```{r scoring}
term_source <- ism_weights |>
  group_by(weight_id, model_family, sample_rep, source_id, term) |>
  summarise(
    score = sum(log1p(9 * value_norm), na.rm = TRUE),
    max_norm = max(value_norm, na.rm = TRUE),
    periods_positive = sum(value_norm > 0, na.rm = TRUE),
    peak_period = period[which.max(value_norm)],
    .groups = "drop"
  ) |>
  group_by(weight_id) |>
  arrange(desc(score), desc(periods_positive), desc(max_norm), term, .by_group = TRUE) |>
  mutate(
    source_rank = row_number(),
    source_score = {
      source_score_max <- max(score, na.rm = TRUE)
      if (is.finite(source_score_max) && source_score_max > 0) {
        score / source_score_max
      } else {
        0
      }
    },
    in_top = source_rank <= 35
  ) |>
  ungroup()

term_summary <- term_source |>
  filter(in_top, source_score > 0) |>
  group_by(term) |>
  summarise(
    n_sources = n_distinct(weight_id),
    n_model_families = n_distinct(model_family),
    n_predictive_families = n_distinct(
      model_family[model_family %in% c("glmnet", "LiblineaR", "naivebayes", "xgboost")]
    ),
    n_reps = n_distinct(sample_rep),
    mean_source_score = mean(source_score, na.rm = TRUE),
    max_source_score = max(source_score, na.rm = TRUE),
    peak_period = round(weighted.mean(peak_period, source_score, na.rm = TRUE) / 10) * 10,
    .groups = "drop"
  ) |>
  arrange(desc(n_model_families), desc(n_sources), desc(mean_source_score), term)

stable_terms <- term_summary |>
  filter(n_model_families >= 3 | n_sources >= 8) |>
  slice_head(n = 45)

selected_terms <- stable_terms$term
```

```{r tbl-stable-terms}
stable_terms |>
  transmute(
    term,
    model_families = n_model_families,
    predictive_families = n_predictive_families,
    sources = n_sources,
    reps = n_reps,
    peak_period,
    mean_score = round(mean_source_score, 3)
  ) |>
  kable(caption = "Most stable ism terms by top-35 support across model/repeat sources.")
```

## 1. Clustered model/repeat support

Each column is one model/repeat source. Each row is an `ism` term. The fill is
the term's normalized score inside that source, so darker cells mean a term was
near the top of that particular model/repeat. Rows and columns are clustered
from the source-by-term matrix.

```{r source-clustering}
source_matrix_df <- term_source |>
  filter(term %in% selected_terms) |>
  select(term, source_id, source_score) |>
  group_by(term, source_id) |>
  summarise(source_score = max(source_score, na.rm = TRUE), .groups = "drop") |>
  complete(term = selected_terms, source_id = info$source_id, fill = list(source_score = 0))

source_wide <- source_matrix_df |>
  pivot_wider(names_from = source_id, values_from = source_score, values_fill = 0) |>
  arrange(term)

source_mat <- as.matrix(source_wide[, -1])
rownames(source_mat) <- source_wide$term

term_order_source <- if (nrow(source_mat) > 1) {
  rownames(source_mat)[hclust(dist(source_mat), method = "ward.D2")$order]
} else {
  rownames(source_mat)
}

source_order <- if (ncol(source_mat) > 1) {
  colnames(source_mat)[hclust(dist(t(source_mat)), method = "ward.D2")$order]
} else {
  colnames(source_mat)
}

source_heat <- source_matrix_df |>
  mutate(
    term = factor(term, levels = rev(term_order_source)),
    source_id = factor(source_id, levels = source_order)
  )
```

```{r fig-clustered-source-heatmap, fig.width=12, fig.height=10}
ggplot(source_heat, aes(source_id, term, fill = source_score)) +
  geom_tile(color = "white", linewidth = 0.15) +
  scale_fill_gradient(low = "#f7fbff", high = "#08306b", limits = c(0, 1)) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Within-source\nscore",
    title = "Clustered support for ism terms across model/repeat sources",
    subtitle = "Rows and columns are clustered; term scores are normalized within each source."
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    axis.text.y = element_text(size = 8),
    panel.grid = element_blank()
  )
```

## 2. Model-family stability

This view collapses repeats within each model family. The fill is how many of
the five repeats place the term in that source's top 35 `ism` terms; the text
shows the same count. Rows are ordered by the weighted peak period so the
historical sequence is visible.

```{r model-family-support}
family_support <- term_source |>
  filter(term %in% selected_terms, in_top, source_score > 0) |>
  group_by(term, model_family) |>
  summarise(
    reps_supported = n_distinct(sample_rep),
    mean_score = mean(source_score, na.rm = TRUE),
    .groups = "drop"
  ) |>
  complete(
    term = selected_terms,
    model_family = factor(model_order, levels = model_order),
    fill = list(reps_supported = 0, mean_score = 0)
  ) |>
  left_join(select(stable_terms, term, peak_period), by = "term")

term_order_period <- stable_terms |>
  arrange(peak_period, term) |>
  pull(term)
```

```{r fig-family-support, fig.width=10, fig.height=9}
ggplot(
  family_support,
  aes(model_family, factor(term, levels = rev(term_order_period)), fill = reps_supported)
) +
  geom_tile(color = "white", linewidth = 0.2) +
  geom_text(aes(label = if_else(reps_supported > 0, as.character(reps_supported), "")), size = 2.6) +
  scale_fill_gradient(low = "#f7fbff", high = "#08519c", limits = c(0, 5), breaks = 0:5) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Repeats\nin top 35",
    title = "Repeat-stable ism terms by model family",
    subtitle = "Rows are ordered by weighted peak decade; numbers count supporting repeats."
  ) +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    axis.text.y = element_text(size = 8),
    panel.grid = element_blank()
  )
```

## 3. Temporal clusters

The previous plots show model/repeat stability. This plot asks when each stable
term is most associated with democracy. Values are the mean normalized positive
association across model/repeat sources in each decade.

```{r temporal-clustering}
term_period <- ism_weights |>
  filter(term %in% selected_terms) |>
  group_by(term, period) |>
  summarise(mean_norm = mean(value_norm, na.rm = TRUE), .groups = "drop") |>
  complete(term = selected_terms, period = sort(unique(ism_weights$period)), fill = list(mean_norm = 0))

period_wide <- term_period |>
  pivot_wider(names_from = period, values_from = mean_norm, values_fill = 0) |>
  arrange(term)

period_mat <- as.matrix(period_wide[, -1])
rownames(period_mat) <- period_wide$term

period_hc <- hclust(dist(period_mat), method = "ward.D2")
period_clusters <- tibble(
  term = names(cutree(period_hc, k = 4)),
  cluster = cutree(period_hc, k = 4)
) |>
  left_join(select(stable_terms, term, peak_period), by = "term") |>
  group_by(cluster) |>
  mutate(cluster_peak = median(peak_period, na.rm = TRUE)) |>
  ungroup() |>
  mutate(
    cluster_order = dense_rank(cluster_peak),
    cluster_label = paste0("Cluster ", cluster_order, " (median peak ", cluster_peak, ")")
  )

term_order_time <- period_clusters |>
  arrange(cluster_order, peak_period, term) |>
  pull(term)

term_period_plot <- term_period |>
  left_join(period_clusters, by = "term") |>
  mutate(term = factor(term, levels = rev(term_order_time)))
```

```{r fig-temporal-clusters, fig.width=11, fig.height=10}
ggplot(term_period_plot, aes(period, term, fill = mean_norm)) +
  geom_tile(color = "white", linewidth = 0.1) +
  facet_grid(cluster_label ~ ., scales = "free_y", space = "free_y") +
  scale_x_continuous(breaks = seq(1700, 2010, 30)) +
  scale_fill_gradient(low = "#fff7ec", high = "#7f0000") +
  labs(
    x = NULL,
    y = NULL,
    fill = "Mean normalized\npositive association",
    title = "Temporal clusters of stable ism terms",
    subtitle = "Values are averaged across model/repeat sources after within-source normalization."
  ) +
  theme(
    axis.text.y = element_text(size = 8),
    strip.text.y = element_text(angle = 0),
    panel.grid = element_blank()
  )
```

## Reading

The most useful terms are those supported by multiple model families and
multiple repeats. Terms that appear in one model family only are better treated
as model-specific diagnostics. In these figures, the stronger stable pattern is
the shift from early constitutional/oppositional `ism` language toward modern
ideological and regime-type language.
