---
title: "Prototype: cross-model similarity & entropy over time"
subtitle: "Democracy full run — figure prototypes for IMPROVEMENT_PLAN W5"
format:
  html:
    toc: true
    toc-location: right
    fig-width: 9
    fig-height: 5
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
theme_set(theme_bw(base_size = 12))

store <- here::here("_targets")
rd <- function(n) targets::tar_read_raw(n, store = store)

ra <- rd("rank_agreements")          # period, term, term2, spearman, top_overlap (+ NA engine cols)
wc <- rd("weight_correlations")      # period, term, term2, value (Pearson)
jsd <- rd("jsd_distances")           # period, term, term2, value (JS distance)
ent <- rd("combined_entropy")        # weight_id, period, entropy, sample_rep, engine, weight_type...

# --- model lookup: weight_id -> family / engine / rep -----------------------
# The similarity targets leave engine/engine2 as NA (a metadata-join bug in
# correlation_functions.R), so we rebuild the mapping from combined_entropy,
# which carries clean labels. Match rate against rank_agreements terms is 100%.
lut <- ent |>
  distinct(weight_id, predictive_model_engine, predictive_model_dfm_weight, weight_type, sample_rep) |>
  mutate(
    family = case_when(
      str_starts(weight_id, "weight_pred") ~ "predictive",
      str_starts(weight_id, "weight_svd")  ~ "svd",
      str_starts(weight_id, "weight_ppmi") ~ "ppmi",
      TRUE ~ "other"
    ),
    model_label = case_when(
      family == "predictive" ~ predictive_model_engine,
      family == "svd"        ~ "PPMI-SVD",
      family == "ppmi" & str_detect(weight_type, "similarities") ~ "PPMI (sim.)",
      family == "ppmi"       ~ "PPMI (feature)",
      TRUE ~ family
    )
  )

engine_pal <- c(glmnet = "#1b9e77", LiblineaR = "#d95f02",
                naivebayes = "#7570b3", xgboost = "#e7298a")
```

## 1. Cross-model weight agreement over time

*How similar are the different classifiers' term-weight rankings within each
decade?* For every decade we take each **rep's** four predictive models
(glmnet, LiblineaR, naïve Bayes, xgboost), compute the Spearman rank
correlation of normalized weights over the common vocabulary for every distinct
engine pair, and average those pairwise agreements. The line is the mean across
the five sampling reps; the band is the rep min–max (replicate range, not a CI —
consistent with Appendix B).

```{r cross-model-prep}
pair <- ra |>
  select(period, term, term2, spearman, top_overlap) |>
  inner_join(select(lut, term = weight_id, fam1 = family, eng1 = model_label, rep1 = sample_rep),
             by = "term") |>
  inner_join(select(lut, term2 = weight_id, fam2 = family, eng2 = model_label, rep2 = sample_rep),
             by = "term2")

# predictive-vs-predictive, same rep, distinct engines
pred_pairs <- pair |>
  filter(fam1 == "predictive", fam2 == "predictive", rep1 == rep2, eng1 != eng2)

per_rep <- pred_pairs |>
  group_by(period, rep = rep1) |>
  summarise(mean_spearman = mean(spearman, na.rm = TRUE),
            mean_overlap  = mean(top_overlap, na.rm = TRUE), .groups = "drop")

agree_summary <- per_rep |>
  group_by(period) |>
  summarise(spearman = mean(mean_spearman),
            lo = min(mean_spearman), hi = max(mean_spearman),
            overlap = mean(mean_overlap), .groups = "drop")
```

```{r fig-cross-model-agreement}
ggplot(agree_summary, aes(period, spearman)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.18, fill = "#1b9e77") +
  geom_line(linewidth = 1) +
  geom_point(size = 1.4) +
  scale_x_continuous(breaks = seq(1700, 2010, 30)) +
  labs(x = NULL, y = "Mean pairwise Spearman\n(4 predictive models)",
       title = "Cross-model agreement in term weights, by decade",
       subtitle = "Line = mean over reps; band = replicate (rep) range") +
  coord_cartesian(ylim = c(0, 1))
```

### Individual engine pairs (rep 1)

Which pairs drive the average? Below, each line is one engine pair in rep 1.

```{r fig-cross-model-pairs}
pairs_r1 <- pred_pairs |>
  filter(rep1 == 1) |>
  mutate(pair = paste(pmin(eng1, eng2), pmax(eng1, eng2), sep = " – ")) |>
  group_by(period, pair) |>
  summarise(spearman = mean(spearman, na.rm = TRUE), .groups = "drop")

ggplot(pairs_r1, aes(period, spearman, colour = pair)) +
  geom_line(linewidth = 0.7) +
  scale_x_continuous(breaks = seq(1700, 2010, 30)) +
  labs(x = NULL, y = "Spearman (rep 1)", colour = "Engine pair",
       title = "Pairwise term-weight agreement between predictive models") +
  coord_cartesian(ylim = c(0, 1)) +
  theme(legend.position = "bottom")
```

### Per-decade agreement heatmap (rep 1, 1950)

```{r fig-cross-model-heatmap, fig.height=4.5, fig.width=6.5}
decade_show <- 1950
hm <- pair |>
  filter(period == decade_show, rep1 == 1, rep2 == 1,
         fam1 %in% c("predictive", "ppmi", "svd"),
         fam2 %in% c("predictive", "ppmi", "svd")) |>
  select(eng1, eng2, spearman)

ggplot(hm, aes(eng1, eng2, fill = spearman)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = sprintf("%.2f", spearman)), size = 3) +
  scale_fill_gradient2(low = "#b2182b", mid = "#f7f7f7", high = "#2166ac",
                       midpoint = 0.5, limits = c(0, 1)) +
  labs(x = NULL, y = NULL, fill = "Spearman",
       title = paste0("Model-vs-model weight agreement, ", decade_show, " (rep 1)")) +
  theme(axis.text.x = element_text(angle = 40, hjust = 1),
        panel.grid = element_blank())
```

## 2. Weight-distribution entropy over time

*How concentrated vs. diffuse is each model's vocabulary weighting?* Shannon
entropy of each model's normalized positive weights per decade. Higher = weight
spread across more words; lower = a few dominant terms. Line = mean over the
five reps, band = rep min–max.

```{r fig-entropy, fig.height=5}
ent_pred <- ent |>
  mutate(family = case_when(
    str_starts(weight_id, "weight_pred") ~ "predictive",
    str_starts(weight_id, "weight_svd")  ~ "svd",
    str_starts(weight_id, "weight_ppmi") ~ "ppmi", TRUE ~ "other")) |>
  filter(family == "predictive") |>
  group_by(engine = predictive_model_engine, period) |>
  summarise(entropy = mean(entropy), lo = min(entropy), hi = max(entropy),
            .groups = "drop")

ggplot(ent_pred, aes(period, entropy, colour = engine, fill = engine)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.15, colour = NA) +
  geom_line(linewidth = 0.9) +
  scale_colour_manual(values = engine_pal, aesthetics = c("colour", "fill")) +
  scale_x_continuous(breaks = seq(1700, 2010, 30)) +
  labs(x = NULL, y = "Entropy (nats) of normalized weights",
       colour = "Model", fill = "Model",
       title = "Vocabulary-weight entropy by predictive model, over time",
       subtitle = "Line = mean over reps; band = replicate (rep) range") +
  theme(legend.position = "bottom")
```

### Entropy including PPMI / SVD baselines

```{r fig-entropy-all, fig.height=5}
ent_all <- ent |>
  left_join(distinct(lut, weight_id, model_label, family), by = "weight_id") |>
  group_by(model_label, period) |>
  summarise(entropy = mean(entropy), lo = min(entropy), hi = max(entropy),
            .groups = "drop")

ggplot(ent_all, aes(period, entropy, colour = model_label, fill = model_label)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.12, colour = NA) +
  geom_line(linewidth = 0.8) +
  scale_x_continuous(breaks = seq(1700, 2010, 30)) +
  labs(x = NULL, y = "Entropy (nats)", colour = "Model", fill = "Model",
       title = "Vocabulary-weight entropy: all model families") +
  theme(legend.position = "bottom")
```

---

*Prototype for W5. Data: `rank_agreements`, `combined_entropy` from the
`full_democracy` run bundle (`_targets/`). Engine labels reconstructed from
`combined_entropy` because the similarity targets currently emit `NA` engine
columns (correlation_functions.R metadata-join bug — flag for W5).*
