---
title: "Prototype: KL divergence & conceptual novelty over time"
subtitle: "Democracy full run — figure prototypes for IMPROVEMENT_PLAN W5"
format:
  html:
    toc: true
    toc-location: right
    fig-width: 9
    fig-height: 6
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(ggrepel)
library(knitr)
theme_set(theme_bw(base_size = 12))

store <- here::here("_targets")
rd <- function(n) targets::tar_read_raw(n, store = store)

model_order <- c(
  "glmnet", "LiblineaR", "naivebayes", "xgboost",
  "PPMI (feature)", "PPMI (sim.)", "PPMI-SVD"
)

# Reuses the sibling prototypes' Dark2-based engine colours; "PPMI (sim.)" is
# swapped from Dark2's brown (too close to "PPMI (feature)" green under
# deuteranopia) to a distinguishable red so the 7-colour set passes the
# categorical CVD-separation check.
model_pal <- c(
  glmnet            = "#1b9e77",
  LiblineaR         = "#d95f02",
  naivebayes        = "#7570b3",
  xgboost           = "#e7298a",
  "PPMI (feature)"  = "#66a61e",
  "PPMI (sim.)"     = "#b2182b",
  "PPMI-SVD"        = "#1f78b4"
)
```

## Aim & data

This prototype builds figures for `combined_kl` (the KL / conceptual-novelty
measures), reading only saved targets from `_targets/`; it does not rerun the
pipeline. It follows the pattern of `prototype_measures.qmd` (cross-model
agreement, entropy) and `prototype_ism_clusters.qmd` (ism-term stability).

`combined_kl` has 527,310 rows and one row per `(P, past_period, future_period,
measure, model source, sampling rep)` combination. Each pairwise KL divergence
between two decades' weight distributions is computed once
(`kl_simple()` in `R/entropy_functions.R`) and then reshaped by
`kl_matrix_to_df()` into three derived measures for every reference decade `P`:

- **novelty** = KL(`P` &#8741; `past_period`), for every `past_period < P` — how much `P`'s
  distribution diverges from an earlier decade.
- **transience** = KL(`P` &#8741; `future_period`), for every `future_period > P` — how much
  `P`'s distribution diverges from a later decade.
- **resonance** = novelty − transience, for every `(past_period, future_period)`
  pair sharing the same `P` (a full cross join within `P`, so most
  `past_period`/`future_period` combinations are *not* adjacent decades — figures
  below filter to the adjacent case, ±1 decade, to get a single well-defined
  resonance value per `P`).

The model/source columns already carry clean labels in `combined_kl` itself
(`predictive_model_engine`, `ppmi_fun`, `svd_dims` are populated directly from
the run's metadata table, unlike `rank_agreements`/`weight_correlations`, which
need the `combined_entropy` lookup fix described in `prototype_measures.qmd`).
There are 7 weight sources (4 predictive engines + PPMI-feature + PPMI-similarity
+ PPMI-SVD) &times; 5 sampling reps &times; 32 decades (1700–2010, 10-year steps), all
using the union vocabulary (`kl_common_vocab` is `FALSE` for every row in this
bundle — the common-vocabulary robustness variant was not run).

```{r data-prep}
kl_raw <- rd("combined_kl")

kl <- kl_raw |>
  transmute(
    P, past_period, future_period, measure, value, sample_rep,
    weight_type, predictive_model_engine, ppmi_fun, svd_dims
  ) |>
  mutate(
    family = case_when(
      !is.na(predictive_model_engine) ~ "predictive",
      !is.na(svd_dims) ~ "svd",
      !is.na(ppmi_fun) ~ "ppmi",
      TRUE ~ "other"
    ),
    model_label = case_when(
      family == "predictive" ~ predictive_model_engine,
      family == "svd" ~ "PPMI-SVD",
      family == "ppmi" & ppmi_fun == "ppmi_similarities" ~ "PPMI (sim.)",
      family == "ppmi" ~ "PPMI (feature)",
      TRUE ~ family
    ),
    model_label = factor(model_label, levels = model_order)
  )

rm(kl_raw)
gc()

diagnostics <- tibble(
  rows = nrow(kl),
  model_sources = n_distinct(kl$model_label),
  reps = n_distinct(kl$sample_rep),
  decades = n_distinct(kl$P),
  measures = n_distinct(kl$measure)
)
```

```{r tbl-diagnostics}
kable(diagnostics, caption = "Data used for the KL prototype.")
```

## 1. Adjacent-decade semantic change

*How much does each decade's weight distribution diverge from the one before
it?* This filters `measure == "novelty"` to `P - past_period == 10` — KL
divergence against the immediately preceding decade only — for every model
source, then averages across the 5 sampling reps. **Panels use free y-axes
deliberately: KL levels are not comparable across model families (a glmnet
weight distribution and an xgboost one are not on the same scale), only the
shape of each model's own trend over time is informative.**

```{r fig1-prep}
adj_novelty <- kl |>
  filter(measure == "novelty", P - past_period == 10) |>
  group_by(model_label, period = P) |>
  summarise(
    mean_kl = mean(value, na.rm = TRUE),
    lo = min(value, na.rm = TRUE),
    hi = max(value, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r fig-adjacent-novelty, fig.width=10, fig.height=7}
ggplot(adj_novelty, aes(period, mean_kl, colour = model_label, fill = model_label)) +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.18, colour = NA) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1) +
  facet_wrap(~model_label, scales = "free_y", ncol = 4) +
  scale_colour_manual(values = model_pal, aesthetics = c("colour", "fill")) +
  scale_x_continuous(breaks = seq(1700, 2010, 60)) +
  labs(
    x = NULL, y = "Adjacent-decade KL divergence (bits)",
    title = "Semantic change: KL divergence from the previous decade",
    subtitle = "Line = mean over 5 sampling reps; band = replicate (rep) range, not a confidence interval"
  ) +
  theme(legend.position = "none", strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 40, hjust = 1, size = 8))
```

*Every predictive engine (glmnet, LiblineaR, naivebayes, xgboost) shows its
highest adjacent-decade novelty in 1710–1750 — glmnet peaks at 1750 (21.6 bits,
mean over reps) and again at 1730 (21.5); LiblineaR and naivebayes peak at 1730
(15.4 and 16.3). This is the part of the corpus where the volume-sampling cap
(500 volumes/decade) is least often reached, so part of the early-decade signal
is corpus scarcity rather than genuine conceptual turnover — treat the
1700–1750 spike as a floor/noise effect, not evidence of unusually fast
semantic change. naivebayes additionally shows a level shift around 1830–1840,
dropping from ~10–16 bits to a stable ~7–9 bits for the rest of the series —
the sharpest structural break of any engine, worth checking against known
corpus-composition changes at that point. glmnet sits on a systematically
higher absolute scale (18–22 bits) than xgboost (3–9 bits) throughout, which is
exactly the "not comparable across engines" caveat in action: LiblineaR and
naivebayes's decade-by-decade shapes correlate at r = 0.96, glmnet correlates
0.86–0.83 with both, but xgboost is essentially uncorrelated with the others
(r = 0.01 with LiblineaR, −0.12 with naivebayes) — its curve looks noisier and
should be read cautiously.*

## 2. Past &times; future KL heatmap

*How far is every decade from every other decade, not just its immediate
neighbours?* `combined_kl` already contains a full novelty/transience
decomposition for one representative engine and rep (glmnet, rep 1): novelty
gives KL(`P` &#8741; `Q`) for every `Q < P` and transience gives it for every
`Q > P`, so combining them reconstructs the complete 32&times;32 directed
divergence matrix (992 off-diagonal cells; the true diagonal, `P == Q`, has KL
= 0 and is dropped upstream).

```{r fig2-prep}
heat_engine <- "glmnet"
heat_rep <- 1

hm <- kl |>
  filter(model_label == heat_engine, sample_rep == heat_rep,
         measure %in% c("novelty", "transience")) |>
  mutate(Q = if_else(measure == "novelty", past_period, future_period)) |>
  filter(!is.na(Q)) |>
  distinct(P, Q, value)
```

```{r fig-kl-heatmap, fig.width=8, fig.height=7}
ggplot(hm, aes(Q, P, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "#f7fbff", high = "#08306b") +
  scale_x_continuous(breaks = seq(1700, 2010, 30)) +
  scale_y_continuous(breaks = seq(1700, 2010, 30)) +
  coord_equal() +
  labs(
    x = "Compared decade (Q)", y = "Reference decade (P)", fill = "KL(P||Q)\n(bits)",
    title = paste0("Past-vs-future KL divergence, ", heat_engine, " (rep ", heat_rep, ")"),
    subtitle = "Cell = KL(P||Q): how much decade P's weight distribution diverges from decade Q's"
  ) +
  theme(panel.grid = element_blank())
```

*The single most divergent pair for glmnet/rep 1 is `P` = 1700 vs `Q` = 2010
(25.6 bits) — the corpus's earliest and latest decades share the least
vocabulary structure, which is the expected result and a sanity check that the
measure behaves as intended. The lowest-divergence pairs are 1720&rarr;1760 and
1880&rarr;1890 (15.6 bits each) — note 1720&rarr;1760 spans 40 years yet is as
close as a true 10-year adjacent pair, a reminder that "adjacent" and "similar"
are not the same thing. The matrix is visibly
asymmetric: KL(1700 &#8741; 2010) = 25.6 but KL(2010 &#8741; 1700) = 24.5 — a modest
but real difference, as expected for a genuinely directional divergence measure
(KL is not a metric). This is one engine and one rep only; it is not averaged
across the replicate range, so read it as illustrative of structure, not as a
precise point estimate.*

## 3. Novelty vs transience (Barron-style resonance)

*Does a decade look more different from its past or from its future?* Both
directions exist in `combined_kl` (confirmed above), so the Barron, Leydesdorff
& Yan (2018) novelty/transience/resonance framing is directly supported.
Resonance is computed here on the adjacent-decade window only (`past_period =
P - 10`, `future_period = P + 10`) so each `(model, rep, decade)` has exactly
one value, matching the framing used in Section 1.

```{r fig3-prep}
adj_resonance <- kl |>
  filter(measure == "resonance", P - past_period == 10, future_period - P == 10) |>
  group_by(model_label, period = P) |>
  summarise(
    mean_res = mean(value, na.rm = TRUE),
    lo = min(value, na.rm = TRUE),
    hi = max(value, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r fig-resonance, fig.width=10, fig.height=6}
ggplot(adj_resonance, aes(period, mean_res, colour = model_label)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1) +
  scale_colour_manual(values = model_pal) +
  scale_x_continuous(breaks = seq(1700, 2010, 30)) +
  labs(
    x = NULL, y = "Resonance = novelty − transience (bits)", colour = "Model",
    title = "Novelty vs transience: adjacent-decade resonance",
    subtitle = "Line = mean over 5 reps. Positive = decade diverges more from its past than its future (novelty-dominant); negative = the reverse (transience-dominant)"
  ) +
  theme(legend.position = "bottom")
```

```{r fig3b-prep}
scatter_engine <- "glmnet"

nov <- kl |>
  filter(model_label == scatter_engine, measure == "novelty", P - past_period == 10) |>
  group_by(period = P) |>
  summarise(novelty = mean(value, na.rm = TRUE), .groups = "drop")

tra <- kl |>
  filter(model_label == scatter_engine, measure == "transience", future_period - P == 10) |>
  group_by(period = P) |>
  summarise(transience = mean(value, na.rm = TRUE), .groups = "drop")

nt <- inner_join(nov, tra, by = "period")
```

```{r fig-novelty-transience-scatter, fig.width=7.5, fig.height=7}
lims <- range(c(nt$novelty, nt$transience))

ggplot(nt, aes(novelty, transience, label = period)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  geom_point(colour = model_pal[[scatter_engine]], size = 2) +
  ggrepel::geom_text_repel(size = 3, max.overlaps = 20) +
  coord_equal(xlim = lims, ylim = lims) +
  labs(
    x = "Novelty: KL(decade || previous decade)", y = "Transience: KL(decade || next decade)",
    title = paste0("Novelty vs transience by decade, ", scatter_engine, " (mean over reps)"),
    subtitle = "Above the diagonal = transience-dominant (looks more different from its future); below = novelty-dominant"
  )
```

*Resonance is small relative to the overall KL scale for every engine — mean
adjacent-decade novelty is 12–19 bits (Section 1) while resonance rarely
exceeds &plusmn;3 bits — so forward/backward asymmetry is a second-order effect on
top of a much larger shared level of decade-to-decade change. The most
transience-dominant point in the whole series is PPMI-SVD at 1820 (resonance =
&minus;1.70: that decade looks more different from 1830 than from 1810); the most
novelty-dominant is xgboost at 1750 (+2.99). Several of the largest resonance
swings cluster in the volume-scarce 1700s–1740s window flagged in Section 1, so
read those specific values with the same corpus-scarcity caveat. For glmnet,
the novelty-vs-transience scatter sits close to the 45&deg; line for most of the
19th–20th century (novelty &#8776; transience — a decade's distance to its past and
future are similar), with the clearest departures at the early, volume-scarce
decades, again reflecting scarcity noise more than a genuine forward/backward
asymmetry in semantic change.*

---

*Prototype for W5. Data: `combined_kl` from the `full_democracy` run bundle
(`_targets/`). Model labels are taken directly from `combined_kl`'s own
metadata columns (`predictive_model_engine`, `ppmi_fun`, `svd_dims`); no
lookup-table repair is needed here (contrast `prototype_measures.qmd`, where
`rank_agreements`/`weight_correlations` need the `combined_entropy` fix for the
correlation_functions.R metadata-join bug). Bands throughout are replicate
(rep) ranges across the 5 sampling reps, not confidence intervals, consistent
with Appendix B's sampling-uncertainty section.*
